Gametic Selection, Sex Ratio Bias, and Transitions Between Sex Determination Systems

### 2016-09-07 LaTeX/Turnover.tex
cleaning up maternal ESD results

\subsection*{Sex chromosome turnover}
The spread of a rare mutant $m$ at the \textbf{M} locus in such a population is determined by the leading eigenvalue, $\lambda$, of the system described by the equations for the next generation frequency of haploid genotypes with the mutation. %, evaluated at the above equilibrium.
Below we present the results for invasion into an $XY$ system assuming no competition among gametes/gametophytes from females and linear arrangement \textbf{X}\textbf{A}\textbf{M}. %equal dominance coefficients in the two sexes,
Equivalent results for invasions into a $ZW$ system are given in the Sup.\ Mat.

@@

%weak selection
Assuming weak selection, we can solve for the invasion fitness of the neo-$Y$ explicitly, giving
\begin{equation}
\lambda_{Y,XY} \approx 1 + V_A \frac{(r - R)s_f}{r R} \frac{s_f}{\left( s_f + s_m \right)^2} t_m^2,
\end{equation}
where $V_A = p_A(1-p_A)$ is the variance at the \textbf{A} locus, $s_f$ and $s_m$ are the respective selection coefficients for $A$ in diploid females and males, $t_m$ is the selection coefficient for $A$ in gametes/gametophytes from males, and we've assumed equal dominance coefficients in the two sexes.
The neo-$Y$ can therefore invade whenever it is in tighter linkage with the selected locus than the ancestral sex-determining locus, $r>R$, provided locus \textbf{A} is polymorphic ($V_A>0$) and there is selection among both females ($s_f\neq0$) and gametes/gametophytes from males ($t_m\neq0$).

@@

%weak selection
Assuming weak selection, we can solve for the invasion fitness of the neo-$W$ explicitly, giving
\begin{equation}
\lambda_{W,XY} \approx 1 + V_A \frac{(2r(1-R)-R) s_f + (1-2r)R s_m }{2 r R} \frac{s_f}{ \left( s_f + s_m \right)^2} t_m^2,
\end{equation}
where we've once again assumed equal dominance coefficients in the two sexes.
In this case, even when the novel sex-determining locus is in looser linkage with the selected locus than is the ancestral sex-determining locus, $r<R$, a novel sex-determiner can invade.
For example, with $R=1/2$ the neo-$W$ invades if there is any linkage between the ancestral sex-determining and selected loci ($r<1/2$), there is selection among gametes/gametophytes in males ($t_m\neq0$), and there is selection for or against $A$ in both males and females ($s_m s_f >0$) that is stronger in females than in males ($|s_f| > |s_m|$).

@@

\begin{equation}
\lambda \approx 1 + V_A C_k B,
\end{equation}
where
%\begin{equation}
%V_A = \frac{- \left[ t_m + h (s_{f}+s_{m}) \right] \left[ t_m + (1-h) (s_{f}+s_{m}) \right]}{(1-2h)^2(s_{f}+s_{m})^2}
%\end{equation}
%is the variance in $A$ ($V_A>0$ when the polymorphic resident equilibrium is stable), 
%with $h$ the dominance coefficient (equal in both sexes), 
\begin{equation}
B = \frac{s_f}{2r R(s_{f}+s_{m})^2} t_m^2
\end{equation}
and $C_k$ is a term that depends on $k$:

%k=0
\begin{equation}
C_0 = 
\end{equation}

%k=1
\begin{equation}
C_1 = - \left[ (2r(1-R)-R) s_f + (1-2r)R s_m \right]
\end{equation}

%k=1/2
\begin{equation}
C_{1/2} = 
\end{equation}

In the case of $k=1$ the maternal and offspring controlled results are the same (which makes sense because we are here concerned only with the passage of mutations through females).

>>>>

\subsection*{Sex chromosome turnover}
The spread of a rare mutant $m$ at the \textbf{M} locus in such a population is determined by the leading eigenvalue, $\lambda$, of the system described by the equations for the next generation frequency of haploid genotypes with the mutation. %, evaluated at the above equilibrium.
Below we present the results for invasion into an $XY$ system assuming no competition among gametes/gametophytes from females and linear arrangement \textbf{X}\textbf{A}\textbf{M} (SAY SOMETHING HERE ABOUT OTHER ARRANGEMENTS). %equal dominance coefficients in the two sexes,
Equivalent results for invasions into a $ZW$ system can be derived by consistently switching the roles of males and females \citep{vanDoorn:2010hu}.

@@

%weak selection
Assuming weak selection, we can solve for the invasion fitness of the neo-$Y$ explicitly, giving
\begin{equation}
\lambda_{Y,XY} \approx 1 + V_A S_A (r - R)s_f,
\end{equation}
where $V_A = p_A(1-p_A)$ is the variance at the \textbf{A} locus, $S_A = s_f/ \left[ r R \left( s_f + s_m \right)^2 t_m^2 \right]$ (INTERPRETATION), $s_f$ and $s_m$ are the respective selection coefficients for $A$ in diploid females and males, $t_m$ is the selection coefficient for $A$ in gametes/gametophytes from males, and we've assumed equal dominance coefficients in the two sexes.
The neo-$Y$ can therefore invade whenever it is in tighter linkage with the selected locus than the ancestral sex-determining locus, $r>R$, provided locus \textbf{A} is polymorphic ($V_A>0$) and there is selection among both females ($s_f\neq0$) and gametes/gametophytes from males ($t_m\neq0$).

@@

%weak selection
Assuming weak selection, we can solve for the invasion fitness of the neo-$W$ explicitly, giving
\begin{equation}
\lambda_{W,XY} \approx 1 + V_A S_A \frac{(2r(1-R)-R) s_f + (1-2r)R s_m }{2},
\end{equation}
where we've once again assumed equal dominance coefficients in the two sexes.
In this case, even when the novel sex-determining locus is in looser linkage with the selected locus than is the ancestral sex-determining locus, $r<R$, a novel sex-determiner can invade.
For example, with $R=1/2$ the neo-$W$ invades if there is any linkage between the ancestral sex-determining and selected loci ($r<1/2$), there is selection among gametes/gametophytes in males ($t_m\neq0$), and there is selection for or against $A$ in both males and females ($s_m s_f >0$) that is stronger in females than in males ($|s_f| > |s_m|$).

@@

\begin{equation}
\lambda \approx 1 + V_A S_A C_k,
\end{equation}
where
%\begin{equation}
%V_A = \frac{- \left[ t_m + h (s_{f}+s_{m}) \right] \left[ t_m + (1-h) (s_{f}+s_{m}) \right]}{(1-2h)^2(s_{f}+s_{m})^2}
%\end{equation}
%is the variance in $A$ ($V_A>0$ when the polymorphic resident equilibrium is stable), 
%with $h$ the dominance coefficient (equal in both sexes), 
$C_k$ is a term that depends on $k$:

%k=0
\begin{equation}
C_0 = big \; ugly \; term
\end{equation}

%k=1
\begin{equation}
C_1 = \frac{\left[ (2r(1-R)-R) s_f + (1-2r)R s_m \right]}{2}
\end{equation}

%k=1/2
\begin{equation}
C_{1/2} = even \; bigger, \; uglier \; term
\end{equation}

In the case of $k=1$ the maternal- and offspring-controlled results are the same since mutation is then only concerned with the passage of mutations through females.
However, with $k\neq0$ the mutant dynamics differ.
